% --------------- 12 POINT FONT -------------------------------
\documentclass[12pt]{article}
% --------------- 10 POINT FONT FOR CAPTIONS ------------------
\usepackage[font=footnotesize]{caption}
% --------------- NY TIMES FONT -------------------------------
\usepackage{times}
% --------------- 1 INCH MARGINS ------------------------------
\usepackage[margin=1in]{geometry}
% --------------- LINE SPACING --------------------------------
\usepackage{setspace}
\singlespacing
%\doublespacing
% --------------- SMALL SECTION TITLES ------------------------
\usepackage[tiny,compact]{titlesec}
% --------------- MATH PACKAGES -------------------------------
\usepackage{amsmath,amsthm,amssymb}
% --------------- Formatting ----------------------------------
%\setlength{\parskip}{\baselineskip}
\setlength{\parindent}{0pt}
\begin{document}
% --------------- TITLE AND NAME ------------------------------
\begin{center}
\underline{\bf Personal Statement - Steven Walton}\\
\end{center}

I grew up reading Science-Fiction and watching Star Trek. This had a lasting
impact on how I view the world, showing how science can change peoples' lives
and the ethical decisions that need to be made. Growing up with role models such
as Isaac Asimov, Carl Sagan, and Captain Kathryn Janeway, I have always been
aware that while science can greatly advance peoples' lives, there are new
ethical dilemmas that arise. As another role model might say: ``With great power
comes great responsibility." As a researcher, I want to act in a way that I
believe would make these role models proud. I believe that this requires me to:
be part teacher, ensuring that my work is accessible and available to others;
part researcher, being creative and performing the best work that I can; and
part philosopher, questioning the impacts and ethical dilemmas of my own
research.

With the rise in popularity of Artificial Intelligence and Machine Learning it
is becoming clear that ethical dilemmas are growing and are frequently being
unanswered, or ignored. Conversely, the rise in AI research has helped push for
more open science, where papers are being published to locations such as ArXiv
and code is being open sourced on platforms such as GitHub. I believe that 
encouraging open science helps everyone, from scientists to the public. 

With open science, I believe the public can be more informed with current
research and help ask and answer the ethical problems that arise. When 
science is locked behind paywalls it causes gate-keeping that only hinders
everyone involved, especially the public and those in impoverished neighborhoods
and countries. Open research leads to higher reproducibility, catching more
mistakes, and allows the public to be involved in the process. I also believe
that when work is performed through public funding that such work should be
accessible to those that funded it.

Having worked on several open sourced scientific projects as well as closed
source, it is apparent to me that open sourced projects are more readily adopted
by the scientific community. It also provides a direct line to the developer and
the users, which leads to higher quality software and a better understanding of
user needs. 

I have always been passionate about learning and trying to understand any
problem that interested me. This lead me to pursue an undergraduate degree in
Space Physics. While I was able to buy all the books that were required for my
classes I found that many times these were not enough. An advantage that I had
was that the material in undergraduate physics textbooks have not change much in
the last hundred years, enabling me to get old text books freely or for very
cheap. I found that sometimes the way one textbook was written would be
difficult for me to understand a topic, but that by going through several
textbooks I could find one that enabled the information to finally sink in. I
found that the more resources available to me, the better I performed. This made
me realize how important it is for experts to write on topics. Though there may
already be existing material that thoroughly explains a subject, sometimes it
requires a different voice for the reader to absorb the knowledge.

Like many other physicists, my colleagues and I admired Richard Feynman. When we
learned about the ``Feynman Technique", which says that one of the best ways to
learn is by teaching, we started to arrange our study groups so that one person
would lecture on a problem and the rest would ask as many questions as they
could. I found this exercise both challenging and exhilarating. Studying by
teaching was a great technique that helped me discover my passion for teaching.

Following through my passion for teaching, I became a Teaching Assistant for
several of the advanced physics labs. I had talked to my professor about my
enthusiasm for teaching and the possibility of becoming a professor after I went
through graduate school. To encourage me, he gave me the opportunity to lecture
several times in his place so that I could get the experience. This directly
lead to me becoming a Supplementary Instructor for a class in Computational
Methods, where I gave a mini lecture twice weekly. I found that this position
challenged me but had great rewards. I loved seeing students succeed and helping
them reach their potential. 

My undergraduate university frequently had outreach days, where we would discuss
with the other colleges and the public about different phenomena in physics. In
my role, I had to design experiments that would both be exciting and could be
discussed in a elegant and understandable manner. Being a Space Physics program,
our outreach frequently revolved around astronomical events, where we would
provide the public and other students the ability to observe events such as
Lunar Eclipses, the Venus Transit, and several comets. I found great joy in
being a science communicator and want that role to always be part of my life. 




%% OLD
%From a young age I have always been interested in the sciences and engineering.
%I grew up on sci-fi books and Star Trek. After I had destroyed a lot of things
%in our house, attempting to figure out how they worked, my dad started handing
%me broken things and challenging me to fix them before he bought replacements. I
%have always had a strong desire to understand how everything works, from the
%microscopic and mundane to the large and exciting areas. 
%
%When I applied to university I was accepted as an aerospace engineer. I wanted
%to explore the stars and thought that this would be the best way to make that
%dream a reality. Before the semester was even over I had changed my degree to
%Physics and working on a research project with the physics chair. I was more
%interested in the toolbox that Physics gives to understand the world than
%strictly building things. In experimental physics I found a nice balance between
%doing hands applications and trying to understand things from a very basic
%level. This passion continued to grow and I started taking on more projects from
%other professors. 
%
%My passion for the stars led to a professor asking me to help
%with running the school's telescope and getting involved in outreach programs.
%Our university would run science events a few times a year for students and
%members of the community. We'd do events mainly around astronomical events like
%solar eclipses, the Venus transit, and lunar eclipses. During the day events we
%would set up a lot of different physics experiments that would get community
%members interested in science and talk about the individual phenomena being
%demonstrated. At night events we would bring out laser pointers and tell some of
%the history and legends surrounding constellations. From these experiences I
%developed a passion for teaching and became a TA for several of the physics
%classes. I took a passion to trying to design lessons that would challenge
%students but also keep them excited and engaged. I also found that his resulted
%in me having a deeper understanding of material and caused me to approach
%problems from different perspectives. 
%
%In undergrad I also got professors to give me extra access to different labs so
%that I could do research on my interests during my free time. Though the
%kindness of these professors I developed a deep passion for research. I felt
%that this access gave me the time to really dig into problems and concentrate on
%solving problems. The free reign gave me the ability to explore topics that I
%would not have been able to during classes and I believe helped develop my
%methods to solving complex problems. 
%
%While in my undergrad I started programming a lot more. I had had a few
%introductions to programming and computer science before, but had never had to
%write anything complicated before. With some of the research the professors had
%given me I had to start writing complicated code and gain a deeper understanding
%of how programming works. At this point I found programming interesting but it
%was a means to an end. 

After graduating I obtained a job in Tennessee to work at a rocket company.
I was able to work on low level research in conjunction with NASA and the
University of Tennessee, Knoxville. While there, I wrote and won a Phase I NASA
Small Business Technology Transfer (STTR) proposal. My employers gave me the
lead position on the project, where I was in charge on planning, running the
experiments, and handling the communication with our University partner
(University of Tennessee, Knoxville). My success in this directly lead to Phase
II funding and a continuation of this work. A large part of this work was
performing computational simulations and then building and testing the resulting
materials that the simulations predicted would have the best performance. This
was my first time working with large and complex simulation code and I found
that I really enjoyed it. I found that programming really resonates with the way
I think and I found it exciting to build simulations that made real world
predictions and that I was making real contributions to science. 

This interest grew to a full on hobby, where I spent a lot of my time outside of
work learning new things about programming and computer science. I found that
there was a large amount of material on the internet that accelerated my
learning. Different blogs, papers, and videos helped me grasp concepts that I
had not learned through my undergraduate classes. I found many exciting topics,
including Artificial Intelligence and Machine Learning, and because of the open
nature of the research I was able to learn a lot about these subjects on my own.

After working for a few years, I decided that I wanted to continue down the path
of the intersection of computer science and the physical sciences. There were
new topics that I had been excited about and teaching myself but felt that this
would be more efficient if I could get a mentor that could help guide me to the
right problems and refine my abilities. I found the High Performance Computing
(HPC) visualization at the University of Oregon (UO) and applied for a PhD
program, where I was accepted and have been a student for a year. 

During my studies at UO, I have been able to work on scientific visualization
projects and contribute to open sourced projects. I believe that my broad set of
skills has added value to our research group, as I can help bridge the
communication between domain scientists and the visualization experts. Through
my current advisor I have had the opportunity to work at several National Labs.
These internships gave me my first on-hands experience with HPC simulations and
the data analysis involved. Working on these projects I saw that there was need
for improvement, especially as there has been a push for larger scale computing.
With my interests in machine learning, physics, mathematics, and computation I
see ample opportunity to combine these topics and that my interdisciplinary
skills would enable me to research this intersection of Machine Learning and
computational analysis.
%While at UO I have been able to work on scientific visualization projects. I
%believe my broad skill set has has been added value to our research group as I
%can help bridge communication between the domain scientists and the
%visualization experts. Through my current advisor I have had the opportunity to
%work at Department of Energy (DoE) National Labs: Oak Ridge National Lab
%(ORNL) and Lawrence Livermore National Lab (LLNL). I got the first internship at
%ORNL the summer before I entered my PhD program. This gave me my first real
%introduction into HPC and how the environment worked. During my second summer I
%was given the opportunity to work at LLNL and given the freedom to pursue my own
%interests. Here I was able to improve both my understanding in machine learning
%and visualizations in HPC. Over the summer I attempted to interpolate scientific
%data through machine learning, the project described within. There were some
%successes and failures in this, but I gained a better understanding of how to
%approach the problem and what tasks need to be solved within. Through my
%experiences here I believe I would be most happy working either in academia or
%at a lab, continuing research and helping mentor new generations of scientists. 

\textbf{Intellectual Merit}
%

With my background in physics and mathematics, coming over to the computer
science field I believe that I am in a unique position for a project such as
this. My strong math and physics background has already enabled me to help my
research group bridge gaps in communication between domain scientists and
visualization experts. It is not uncommon for miscommunication to happen between
different research domains, as researchers may be using similar words but in
different contexts. This has helped our group better understand what domain
scientists are looking for, but may have a difficult time communicating. Having
come from a background of performing simulations, it is easier to intuit what a
simulation scientist is trying to accomplish.

My strong background in mathematics has enabled me to quickly catch up with
modern machine learning techniques and understand the underlying statistical
properties of models. Being in a computer science environment has helped reduce
my so called ``Swiss Cheese" knowledge in computer science and HPC and allow me 
to progress towards expertise. 

\textbf{Broader Impacts}
%

I believe that one of the most important duties of a researcher is being able to
communicate their work. Scientists such as Carl Sagan and Neil deGrasse Tyson
have impacted my life and I believe their method of communication should be
emulated by the scientific community. I have greatly enjoyed my experiences
teaching classes and would like to become a professor in the future. I believe
that becoming a professor would allow me to follow through with my passion for
research and teaching. 

Additionally, I am passionate about open source software and blogging.
Personally, I have benefited greatly from both of these and blogging has enabled
me to learn a large breadth of subjects. In an effort to contribute back to the
community, which has helped me so much, I try to open source as much software as
I can and write in my blog about what I am learning. I believe that a NSF
fellowship puts me in a position that will encourage these passions. Not only
will it free up time to enable more blogging of my research, but through a
public funding I am not restricted from open sourcing my research and having the
research be in the open.

I believe that by open sourcing and blogging about my research that this can
more greatly enable reproducibility of results and push science to be more open.
It is my belief that science is performed more efficiently when more eyes are
able to see what is being done and more people are able to perform research.
With funding through the NSF, I am able to perform research without constraints
of visibility and can perform my research in the open.

%I not only have experience in teaching, but have enjoyed the opportunities that
%I have had to do so. An aspect that I greatly appreciate about my current
%research group is that there is a large focus on communication skills. Not only
%skills of how to communicate ideas between scientists in different domains, but
%to larger audiences. I believe that it is becoming increasingly important that
%all scientists become better communicators. Specifically I think we need to
%learn how to better explain our research to the general public and the
%importance of that work. Not only do I think it helps us get funding for our
%work, but helps get more people into science, and helps the world become more
%scientifically friendly. 
%
%I also think it is important for science to be done in the open, especially when
%it is publicly funded. In this effort I try to write about my own work and open
%source it through online media such as GitHub. I believe open source is both
%good for the scientific community as well as the public. It helps the scientific
%community reproduce and check work by other scientists. It also helps the public
%by making the science available to everyone. 

\end{document}

